---
title: "ClimateWMA"
author: "KateMarkham"
date: "06/08/2022"
output:
  html_document: default
---
This script plots climatic and environmental variables for Hwange District in Zimbabwe.  Data is then plotted, correlation is checked, and tests for normality are done. To date, data has not fit normal distribution. Thus, the Kruskal-Wallis test is used to determine differences. Following any statistically significant differences, the Dunn's Kruskal-Wallis multiple comparisons is to identify directional differences and if they are significant. 

This script was designed to take csv files as inputs. Each csv file contains measures of central tendency by month of for either land cover type or by communal area. Climatic and environmental variables included are LST, soil moisture, precipitation, and NDVI. Separate inputs are needed for each environmental variable (LST, NDVI, etc.) and for communal land versus land cover type.

Part I examines climatic and environmental variables by communal lands (also called wildlife management areas).
Part II examines the same as above but for the crop raiding season only (March, April, and May).
Part III examines variables by land cover type.
Part IV examines the same as above but for the crop raiding season only.




Load packages needed
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(data.table)
library(readr)
library(ggplot2)
library(gridExtra)
library(car)
library(stargazer)
library(tmap)
library(lmtest)
library(sf)
library(Hmisc)
library(FSA)
```

***I. Communal lands-all months***

Import all of the communal lands data and make Year and Month factors rather than numeric.
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Import data
WMA_LST <- read_csv("WMA_LST.csv")
WMA_LST$Year<-factor(WMA_LST$Year)
WMA_LST$Month<-factor(WMA_LST$Month)
LST<-WMA_LST

WMA_NDVI <- read_csv("WMA_NDVI.csv")
WMA_NDVI$Year<-factor(WMA_NDVI$Year)
WMA_NDVI$Month<-factor(WMA_NDVI$Month)
NDVI<-WMA_NDVI 

WMA_Precip <- read_csv("WMA_Precip.csv")
WMA_Precip$Year<-factor(WMA_Precip$Year)
WMA_Precip$Month<-factor(WMA_Precip$Month)
Precip<-WMA_Precip
  
WMA_Soil <- read_delim("WMA_Soil.csv", delim = "\t", 
    escape_double = FALSE, trim_ws = TRUE)
WMA_Soil$Year<-factor(WMA_Soil$Year)
WMA_Soil$Month<-factor(WMA_Soil$Month)
Soil<-WMA_Soil
```


Plot LST in communal lands
```{r}

ggplot(LST,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median LST in communal lands")+ 
  xlab("Month") + ylab("LST (K)") +
  facet_wrap(~Year)
```


Plot precip by year in all communal lands
```{r}

ggplot(Precip,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median precipitation in communal lands")+ 
  xlab("Month") + ylab("Precipitation (mm/month)") +
  facet_wrap(~Year)
```

Soil moisture 10 to 40 cm by year in all communal lands
```{r}

ggplot(Soil,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median soil moisture (10 to 40 cm depth) in communal lands")+ 
  xlab("Month") + ylab("Soil moisture (m3/m3)") +
  facet_wrap(~Year)

ggplot(Soil,aes(x=Month,y=RANGE))+
  geom_boxplot() + ggtitle("Soil moisture (10 to 40 cm depth) range in communal lands")+ 
  xlab("Month") + ylab("Soil moisture (m3/m3)") +
  facet_wrap(~Year)

ggplot(Soil,aes(x=Month,y=MAX))+
  geom_boxplot() + ggtitle("Maximum soil moisture (10 to 40 cm depth) in communal lands")+ 
  xlab("Month") + ylab("Soil moisture (m3/m3)") +
  facet_wrap(~Year)

```

NDVI by year in all communal lands
```{r}

ggplot(NDVI,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median NDVI in communal lands")+ 
  xlab("Month") + ylab("Median NDVI (1-200)") +
  facet_wrap(~Year)


ggplot(NDVI,aes(x=Month,y=RANGE))+
  geom_boxplot() + ggtitle("NDVI Range in communal lands")+ 
  xlab("Month") + ylab("NDVI Range (1-200)") +
  facet_wrap(~Year)
```



Test for normality
Shapiro test: if p<0.01, data is not normally distributed.
```{r}
shapiro.test(NDVI$MEDIAN) 
qqnorm(NDVI$MEDIAN); qqline(NDVI$MEDIAN)
```
Test and plot differences in NDVI by WMA
```{r warning=FALSE}

##only data for which we have CR reports:
#WMA_NDVI<- WMA_NDVI %>%
#  filter(NAME%in% c("Chidobe", "Jambezi", "Kachecheti","Nemananga"))
kruskal.test(MEDIAN ~ NAME, data = WMA_NDVI) 
kruskal.test(RANGE ~ NAME, data = WMA_NDVI) 
dunnTestWMA_NDVIRANGE<-dunnTest(RANGE ~ NAME, data = WMA_NDVI, method = "bh",list=TRUE)
dunnTestWMA_NDVIRANGE

kruskal.test(MAX ~ NAME, data = WMA_NDVI) 

# ggplot(WMA_NDVI,aes(x=NAME,y=MEDIAN))+
#   geom_boxplot() + ggtitle("NDVI (0-200) in communal lands")+ 
#   xlab("Communal land") + ylab("NDVI") +
#   facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Tests and plots for precip and LST by communal land
```{r}
kruskal.test(MEDIAN ~ NAME, data = WMA_Precip) 
kruskal.test(RANGE ~ NAME, data = WMA_Precip) 
kruskal.test(MAX ~ NAME, data = WMA_Precip) 

kruskal.test(MEDIAN ~ NAME, data = WMA_LST) 
kruskal.test(RANGE ~ NAME, data = WMA_LST) 
kruskal.test(MAX ~ NAME, data = WMA_LST) 
```


Tests and plots for soil moisture distribution by communal land
```{r}

kruskal.test(MEDIAN ~ NAME, data = WMA_Soil) 
dunnTestWMA_SoilMEDIAN<-dunnTest(MEDIAN ~ NAME, data = WMA_Soil, method = "bh",list=TRUE)
dunnTestWMA_SoilMEDIAN
kruskal.test(RANGE ~ NAME, data = WMA_Soil) 
kruskal.test(MAX ~ NAME, data = WMA_Soil) 

# ggplot(WMA_Soil,aes(x=NAME,y=MEDIAN))+
#   geom_boxplot() + ggtitle("Median soil moisture (10 to 40 cm depth) in communal lands")+ 
#   xlab("Communal land") + ylab("Soil moisture (m3/m3)") +
#   facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# ggplot(WMA_Soil,aes(x=NAME,y=RANGE))+
#   geom_boxplot() + ggtitle("Soil moisture (10 to 40 cm depth) range in communal lands")+ 
#   xlab("Communal land") + ylab("Soil moisture (m3/m3)") +
#   facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# ggplot(WMA_Soil,aes(x=NAME,y=MAX))+
#   geom_boxplot() + ggtitle("Maximum soil moisture (10 to 40 cm depth) in communal lands")+ 
#   xlab("Communal land") + ylab("Soil moisture (m3/m3)") +
#   facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

******II. Communal lands-crop raiding only******


Crop raiding months only
LST WMA
```{r echo=TRUE, message=FALSE, warning=FALSE}

WMA_LST_CR<-WMA_LST%>% 
  filter(Month%in% c("03", "04", "05")) %>%
  filter(NAME%in% c("Chidobe", "Jambezi", "Kachecheti","Nemananga"))

# WMA_CR_LST_Plot<-ggplot(WMA_LST_CR,aes(x=Month,y=MEDIAN, color=NAME))+
#   geom_point() + ggtitle("Median LST in communal lands during crop raiding season")+ 
#   xlab("Month") + ylab("LST (K)") +
#   facet_wrap(~Year)
# WMA_CR_LST_Plot1<-WMA_CR_LST_Plot + scale_color_discrete(name="Communal land")
# WMA_CR_LST_Plot
```

Precipitation three months prior to CR period due to assumed lag WMA
```{r echo=TRUE, message=FALSE, warning=FALSE}

WMA_Precip_CR<-WMA_Precip %>% 
  filter(Month%in% c("12", "01", "02"))

# WMA_CR_Precip<-ggplot(WMA_Precip_CR,aes(x=Month,y=MEDIAN, color=NAME))+
#   geom_point() + ggtitle("Median precipitation in communal lands during crop raiding season")+ 
#   xlab("Month") + ylab("Precipitation (mm/month)") +
#   facet_wrap(~Year)
# WMA_CR_Precip1<-WMA_CR_Precip + scale_color_discrete(name="Communal land")
# WMA_CR_Precip1
```

Soil moisture WMA crop raiding period
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Soil Moisture Plots for crop raiding period 
#Median NDVI
WMA_Soil_CR<-WMA_Soil %>% 
  filter(Month %in% c(3, 4, 5)) %>%
  filter(NAME%in% c("Chidobe", "Jambezi", "Kachecheti","Nemananga"))
# p1<-ggplot(WMA_Soil_CR,aes(x=Month,y=MEDIAN,color=NAME))+
#   geom_point()+
#   xlab("Month") + ylab("Median Soil Moisture (m3/m3)") +ggtitle("Median soil mosture in communal lands during crop raiding season") +
#   facet_wrap(~Year)
# p1<-p1 + scale_color_discrete(name="Communal land")
# p1


kruskal.test(MEDIAN ~ NAME, data = WMA_Soil_CR) 
dunnTestWMA_NDVI_CR<-dunnTest(MAX ~ NAME, data = WMA_Soil_CR, method = "bh",list=TRUE)
```

NDVI WMA CROP RAIDING PERIOD
```{r echo=TRUE, message=FALSE, warning=FALSE}
#NDVI Plots for crop raiding period 
#MedianNDVI
WMA_NDVI_CR<-WMA_NDVI %>% 
  filter(Month%in% c("03", "04", "05")) %>%
  filter(NAME%in% c("Chidobe", "Jambezi", "Kachecheti","Nemananga"))
# p1<-ggplot(WMA_NDVI_CR,aes(x=Month,y=MEDIAN,color=NAME))+
#   geom_point()+
#   xlab("Month") + ylab("Median NDVI (0-200)") +ggtitle("Median NDVI in communal lands during crop raiding season") +
#   facet_wrap(~Year)
# p1<-p1 + scale_color_discrete(name="Communal land")
# p1

kruskal.test(STD ~ NAME, data = WMA_NDVI_CR) 
dunnTestWMA_NDVI_CR<-dunnTest(STD ~ NAME, data = WMA_NDVI_CR, method = "bh",list=TRUE)
dunnTestWMA_NDVI_CR
```


***III. Land Cover Analysis-all months***
Import data for land cover
```{r echo=TRUE, message=FALSE, warning=FALSE}
LC_LST <- read_csv("LC_LST.csv")
LC_LST$Year<-factor(LC_LST$Year)
LC_LST$Month<-factor(LC_LST$Month)
LC_LST<-LC_LST %>% 
  filter(LC %in% c("ag", "forest", "grassland","shrubland"))

LC_Precip <- read_csv("LC_Precip.csv")
LC_Precip$Year<-factor(LC_Precip$Year)
LC_Precip$Month<-factor(LC_Precip$Month)
LC_Precip <-LC_Precip %>% 
  filter(LC %in% c("ag", "forest", "grassland","shrubland"))

LC_Soil<-LC_Soil <- read_csv("LC_Soil.csv")
LC_Soil$Year<-factor(LC_Soil$Year)
LC_Soil$Month<-factor(LC_Soil$Month)
LC_Soil<-LC_Soil %>% 
  filter(LC %in% c("ag", "forest", "grassland","shrubland"))

LC_NDVI <- read_csv("LC_NDVI.csv")
LC_NDVI$Year<-factor(LC_NDVI$Year)
LC_NDVI$Month<-factor(LC_NDVI$Month)
LC_NDVI<-LC_NDVI %>% 
  filter(LC %in% c("ag", "forest", "grassland","shrubland"))
```


LST 
```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(LC_LST,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median LST in different land cover types")+ 
  xlab("Month") + ylab("LST (K)") +
  facet_wrap(~Year)

LST_LC_Plot<-ggplot(LC_LST,aes(x=Month,y=MEDIAN, color=LC))+
  geom_point() + ggtitle("Median LST by Land Cover")+ 
  xlab("Month") + ylab("LST (K)") +
  facet_wrap(~Year)
LST_LC_Plot1<-LST_LC_Plot + scale_color_discrete(name="LC")
LST_LC_Plot1
```

Precip 
```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(LC_Precip,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median Precip in different land cover types")+ 
  xlab("Month") + ylab("Precipitation (mm)") +
  facet_wrap(~Year)

Precip_LC_Plot<-ggplot(LC_Precip,aes(x=Month,y=MEDIAN, color=LC))+
  geom_point() + ggtitle("Median Precipitation by Land Cover")+ 
  xlab("Month") + ylab("Precipitation (mm)") +
  facet_wrap(~Year)
Precip_LC_Plot1<-Precip_LC_Plot + scale_color_discrete(name="LC")
Precip_LC_Plot1
```


NDVI 
```{r echo=TRUE, message=FALSE, warning=FALSE}
ggplot(LC_NDVI,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median NDVI in different land cover types")+ 
  xlab("Month") + ylab("NDVI") +
  facet_wrap(~Year)

LC_NDVI_Plot<-ggplot(LC_NDVI,aes(x=Month,y=MEDIAN, color=LC))+
  geom_point() + ggtitle("Median NDVI by Land Cover")+ 
  xlab("Month") + ylab("NDVI (0-200)") +
  facet_wrap(~Year)
LC_NDVI_Plot1<-LC_NDVI_Plot + scale_color_discrete(name="LC")
LC_NDVI_Plot1
```



Soil moisture
```{r}
ggplot(LC_Soil,aes(x=Month,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median Soil Moisture in different land cover types")+ 
  xlab("Month") + ylab("Soil Moisture (m3/m3)") +
  facet_wrap(~Year)

LC_Soil_Plot<-ggplot(LC_Soil,aes(x=Month,y=MEDIAN, color=LC))+
  geom_point() + ggtitle("Median Soil Moisture by Land Cover")+ 
  xlab("Month") + ylab("Soil Moisture (m3/m3)") +
  facet_wrap(~Year)
LC_Soil_Plot1<-LC_Soil_Plot + scale_color_discrete(name="LC")
LC_Soil_Plot1
```


Test for normality
Shapiro test: if p<0.01, data is not normally distributed.
```{r}
shapiro.test(LC_NDVI$MEDIAN) 
qqnorm(LC_NDVI$MEDIAN); qqline(LC_NDVI$MEDIAN)

shapiro.test(LC_Precip$MEDIAN) 
qqnorm(LC_Precip$MEDIAN); qqline(LC_Precip$MEDIAN)

shapiro.test(LC_Soil$MEDIAN) 
qqnorm(LC_Soil$MEDIAN); qqline(LC_Soil$MEDIAN)

shapiro.test(LC_LST$MEDIAN) 
qqnorm(LC_LST$MEDIAN); qqline(LC_LST$MEDIAN)
```
Data is not normally distributed.


Shapiro test: if p<0.01, data is not normally distributed.
```{r}
shapiro.test(LC_NDVI$PCT90) 
qqnorm(LC_NDVI$PCT90); qqline(LC_NDVI$PCT90)

shapiro.test(LC_Precip$PCT90) 
qqnorm(LC_Precip$PCT90); qqline(LC_Precip$PCT90)

shapiro.test(LC_Soil$PCT90) 
qqnorm(LC_Soil$PCT90); qqline(LC_Soil$PCT90)

shapiro.test(LC_LST$PCT90) 
qqnorm(LC_LST$PCT90); qqline(LC_LST$PCT90)
```

Shapiro test: if p<0.01, data is not normally distributed.
```{r}
shapiro.test(LC_NDVI$RANGE) 
qqnorm(LC_NDVI$RANGE); qqline(LC_NDVI$RANGE)

shapiro.test(LC_Precip$RANGE) 
qqnorm(LC_Precip$RANGE); qqline(LC_Precip$RANGE)

shapiro.test(LC_Soil$RANGE) 
qqnorm(LC_Soil$RANGE); qqline(LC_Soil$RANGE)

shapiro.test(LC_LST$RANGE) 
qqnorm(LC_LST$RANGE); qqline(LC_LST$RANGE)
```
Test for correlation with variables of interest and area to check that there is nothing weird happening with large pixel size
```{r}
cor(LC_Soil$MEDIAN, LC_Soil$AREA)
cor(LC_NDVI$MEDIAN,LC_NDVI$AREA)
cor(LC_Precip$MEDIAN,LC_Precip$AREA)
cor(LC_LST$MEDIAN,LC_LST$AREA)

cor(LC_Soil$MAX, LC_Soil$AREA)
cor(LC_NDVI$MAX,LC_NDVI$AREA)
cor(LC_Precip$MAX,LC_Precip$AREA)
cor(LC_LST$MAX,LC_LST$AREA)

cor(LC_Soil$RANGE, LC_Soil$AREA)
cor(LC_NDVI$RANGE,LC_NDVI$AREA)
cor(LC_Precip$RANGE,LC_Precip$AREA)
cor(LC_LST$RANGE,LC_LST$AREA)
```

**Statistics**
```{r warning=FALSE}

kruskal.test(MEDIAN ~ LC, data = LC_NDVI) 
dunnTestLC_NDVIMEDIAN<-dunnTest(MEDIAN ~ LC, data = LC_NDVI, method = "bh",list=TRUE)
dunnTestLC_NDVIMEDIAN
kruskal.test(RANGE ~ LC, data = LC_NDVI) 
dunnTestLC_NDVIRANGE<-dunnTest(RANGE ~ LC, data = LC_NDVI, method = "bh",list=TRUE)
dunnTestLC_NDVIRANGE

kruskal.test(MAX ~ LC, data = LC_NDVI) 
dunnTestLC_NDVIMAX <-dunnTest(MAX ~ LC, data = LC_NDVI, method = "bh",list=TRUE)
dunnTestLC_NDVIMAX 

ggplot(LC_NDVI,aes(x=LC,y=MEDIAN))+
  geom_boxplot() + ggtitle("Median NDVI (0-200) by land cover type")+ 
  xlab("Land cover") + ylab("NDVI") +
  facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(LC_NDVI,aes(x=LC,y=RANGE))+
  geom_boxplot() + ggtitle("Range in NDVI (0-200) by land cover type")+ 
  xlab("Land cover") + ylab("NDVI") +
  facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
Precipitation
```{r warning=FALSE}
kruskal.test(MEDIAN ~ LC, data = LC_Precip) 
dunnTestLC_PrecipMEDIAN<-dunnTest(MEDIAN ~ LC, data = LC_Precip, method = "bh",list=TRUE)
dunnTestLC_PrecipMEDIAN
kruskal.test(RANGE ~ LC, data = LC_Precip) 
dunnTestLC_PrecipRANGE<-dunnTest(RANGE ~ LC, data = LC_Precip, method = "bh",list=TRUE)
dunnTestLC_PrecipRANGE

kruskal.test(MAX ~ LC, data = LC_Precip) 
dunnTestLC_PrecipMAX<-dunnTest(MAX ~ LC, data = LC_Precip, method = "bh",list=TRUE)
dunnTestLC_PrecipMAX

ggplot(LC_Precip,aes(x=LC,y=MEDIAN))+
  geom_boxplot() + ggtitle("Precipitation  by land cover type")+ 
  xlab("Land cover") + ylab("Precipitation (mm)") +
  facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(LC_Precip,aes(x=LC,y=RANGE))+
  geom_boxplot() + ggtitle("Range in Precipitation by land cover type")+ 
  xlab("Land cover") + ylab("Precipitation (mm)") +
  facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
Soil moisture
```{r warning=FALSE}
kruskal.test(MEDIAN ~ LC, data = LC_Soil) 

kruskal.test(RANGE ~ LC, data = LC_Soil) 
dunnTestLC_SoilRANGE<-dunnTest(RANGE ~ LC, data = LC_Soil, method = "bh",list=TRUE)
dunnTestLC_SoilRANGE

kruskal.test(MAX ~ LC, data = LC_Soil) 
dunnTestLC_SoilMAX<-dunnTest(MAX ~ LC, data = LC_Soil, method = "bh",list=TRUE)
dunnTestLC_SoilMAX

ggplot(LC_Soil,aes(x=LC,y=MAX))+
  geom_boxplot() + ggtitle("Maximum soil moisture at 10-40 cm by land cover type")+ 
  xlab("Land cover") + ylab("Soil moisture (m3/m3)") +
  facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(LC_Soil,aes(x=LC,y=RANGE))+
  geom_boxplot() + ggtitle("Range in soil moisture at 10-40 cm by land cover type")+ 
  xlab("Land cover") + ylab("Soil moisture (m3/m3)") +
  facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
LST

```{r warning=FALSE}
kruskal.test(MEDIAN ~ LC, data = LC_LST) 

kruskal.test(RANGE ~ LC, data = LC_LST) 
dunnTestLSTRANGE<-dunnTest(RANGE ~ LC, data = LC_LST, method = "bh",list=TRUE)
dunnTestLSTRANGE

kruskal.test(MAX ~ LC, data = LC_LST) 




ggplot(LC_LST,aes(x=LC,y=RANGE))+
  geom_boxplot() + ggtitle("Range in LST by land cover type")+ 
  xlab("Land cover") + ylab("LST (K)") +
  facet_wrap(~Year) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


***IV. Land Cover-crop raiding season***

```{r echo=TRUE, message=FALSE, warning=FALSE}
#NDVI Plots for crop raiding period 
#Range NDVI
LC_NDVI_CR<-LC_NDVI %>% 
  filter(Month%in% c("03", "04", "05"))
p1<-ggplot(LC_NDVI_CR,aes(x=Month,y=MEDIAN,color=LC))+
  geom_point()+
  xlab("Month") + ylab("Median NDVI (0-200)") +ggtitle("Median NDVI by land cover during crop raiding season") +
  facet_wrap(~Year)
p1<-p1 + scale_color_discrete(name="Land cover")
p1
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
#NDVI Plots for crop raiding period 
#Range NDVI
LC_NDVI_CR<-LC_NDVI %>% 
  filter(Month%in% c("03", "04", "05"))
p1<-ggplot(LC_NDVI_CR,aes(x=Month,y=RANGE,color=LC))+
  geom_point()+
  xlab("Month") + ylab("Range NDVI (0-200)") +ggtitle("Range in NDVI values by land cover during crop raiding season") +
  facet_wrap(~Year)
p1<-p1 + scale_color_discrete(name="Land cover")
p1
```

Maximum NDVI CR by LC
```{r echo=TRUE, message=FALSE, warning=FALSE}
#NDVI Plots for crop raiding period 
#Maximum NDVI
LCNDVI_CR<-LC_NDVI %>% 
  filter(Month%in% c("03", "04", "05"))
p1<-ggplot(LCNDVI_CR,aes(x=Month,y=MAX,color=LC))+
  geom_point()+
  xlab("Month") + ylab("Maximum NDVI (0-200)") +ggtitle("Maximum NDVI during crop raiding season by land cover") +
  facet_wrap(~Year)
p1<-p1 + scale_color_discrete(name="Land cover type")
p1
```

Statistics CR period only-NDVI

```{r}
kruskal.test(MEDIAN ~ LC, data = LCNDVI_CR) 
dunnTestLCNDVI_CR_MEDIAN<-dunnTest(MEDIAN ~ LC, data = LCNDVI_CR, method = "bh",list=TRUE)
dunnTestLCNDVI_CR_MEDIAN

kruskal.test(RANGE ~ LC, data = LCNDVI_CR) 
dunnTestLCNDVI_CR_RANGE<-dunnTest(RANGE ~ LC, data = LCNDVI_CR, method = "bh",list=TRUE)
dunnTestLCNDVI_CR_RANGE

kruskal.test(MAX ~ LC, data = LCNDVI_CR) 
dunnTestLCNDVI_CR_MAX<-dunnTest(MAX ~ LC, data = LCNDVI_CR, method = "bh",list=TRUE)
dunnTestLCNDVI_CR_MAX
```
Statistics precipitation CR period only
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Precip Plots for crop raiding period 
#Range Precip
LC_Precip_CR<-LC_Precip %>% 
  filter(Month%in% c("03", "04", "05"))
p1<-ggplot(LC_Precip_CR,aes(x=Month,y=RANGE,color=LC))+
  geom_point()+
  xlab("Month") + ylab("Precipitation Range (mm)") +ggtitle("Precipitation by land cover prior to crop raiding season") +
  facet_wrap(~Year)
p1<-p1 + scale_color_discrete(name="Land cover")
p1
```
Statistics CR period only-precip

```{r}
kruskal.test(MEDIAN ~ LC, data = LC_Precip_CR) 
dunnTestLCPrecip_CR_MEDIAN<-dunnTest(MEDIAN ~ LC, data = LC_Precip_CR, method = "bh",list=TRUE)
kruskal.test(RANGE ~ LC, data = LC_Precip_CR) 
dunnTestLCPrecip_CR_RANGE<-dunnTest(RANGE ~ LC, data = LC_Precip_CR, method = "bh",list=TRUE)
kruskal.test(MAX ~ LC, data = LC_Precip_CR) 
dunnTestLCPrecip_CR_MAX<-dunnTest(MAX ~ LC, data = LC_Precip_CR, method = "bh",list=TRUE)
```

Statistics CR period only-soil

```{r echo=TRUE, message=FALSE, warning=FALSE}
#Soil Plots for crop raiding period 
LC_Soil_CR<-LC_Soil %>% 
  filter(Month%in% c("3", "4", "5"))
         
kruskal.test(MEDIAN ~ LC, data = LC_Soil_CR) 
kruskal.test(RANGE ~ LC, data = LC_Soil_CR) 
dunnTestLCSoil_CR_RANGE<-dunnTest(RANGE ~ LC, data = LC_Soil_CR, method = "bh",list=TRUE)
dunnTestLCSoil_CR_RANGE
kruskal.test(MAX ~ LC, data = LC_Soil_CR) 
dunnTestLCSoil_CR_MAX<-dunnTest(MAX ~ LC, data = LC_Soil_CR, method = "bh",list=TRUE)
dunnTestLCSoil_CR_MAX
```


Stats CR period only-LST
```{r echo=TRUE, message=FALSE, warning=FALSE}
#Soil Plots for crop raiding period 
#Range soil
LC_LST_CR<-LC_LST %>% 
  filter(Month%in% c("03", "04", "05"))
kruskal.test(MEDIAN ~ LC, data = LC_LST_CR) 
kruskal.test(RANGE ~ LC, data = LC_LST_CR) 
kruskal.test(MAX ~ LC, data = LC_LST_CR) 
dunnTestLCLST_CR_RANGE<-dunnTest(RANGE ~ LC, data = LC_LST_CR, method = "bh",list=TRUE)
dunnTestLCLST_CR_RANGE

```